---
title: "Determining City Type"
author: "Olivia Beck"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(compensator)
```

To determine which city type a nonprofit is located in, we use county-level Federal Information Processing System (FIPS) codes and Rural-Urban Commuting Area (RUCA) codes.

According the to Federal Communications Commission:

> "FIPS codes are numbers which uniquely identify geographic areas. The number of digits in FIPS codes vary depending on the level of geography. State-level FIPS codes have two digits, county-level FIPS codes have five digits of which the first two are the FIPS code of the state to which the county belongs"

A list of all available FIPS codes can be found [here](https://transition.fcc.gov/oet/info/maps/census/fips/fips.txt). When the IRS publishes tax return data for nonprofits, they include the 5-digit ZIP code of the primary address listed on Form 990. We can easily find the FIPS code for each nonprofit using their respective ZIP code by utalizing the cross walk published by the U.S. Department of Housing and Urban Development (HUD) [here](https://www.huduser.gov/portal/datasets/usps.html).

In order to obtain the city type from the FIPS code, we use the RUCA codes from the U.S. Department of Agriculture (USDA). RUCA codes assign each county a score of 1 - 10 from most metropolitan to most rural based on the largest commuting flow in the county. For our purposes we assign RUCA code to be metropolitan, codes 2- 4 to be suburban, 4-6 as small town, and 7-10 as rural. A crosswalk between county-level FIPS codes and RUCA codes can be found on the USDA's website [here](https://www.ers.usda.gov/data-products/rural-urban-commuting-area-codes/). As of this date, the USDA is currently using the 2010 Census to determine the RUCA codes, but has plans to update this data with the 2020 Census soon. We are currently using the 2010 Census based RUCA codes.

### Examples 

We provide a few examples of assigning a city location to a nonprofit.

The Bucks County Historical Society operates a few small museums in Bucks County, Pennsylvania (located between Trenton, Allentown, and Philadelphia). Their EIN is 231371952 and they report their ZIP code as 18901. We use the HUD ZIP-to-FIPS cross walk to find the FIPS code for 18901 is 42017. We then use the USDA's FIPS-to-RUCA crosswalk to find that 42017 has a RUCA code of 1. We conclude Bucks County Historical Society is located in a metropolitan city.

```{r}
nonprofits[nonprofits$EIN == "231371952", c("EIN", "name", "state", "zip5")]

```

The Arkansas Methodist Medical Center is a hospital and health clinic located in Paragould, Arkansas. Their EIN is 721562468 and they report their ZIP code as 72450, which is located in Greene County. Greene County, Arkansas has FIPS code 05055, which has RUCA code 5, so we conclude The Arkansas Methodist Medical Center is located in a small town.

```{r}
nonprofits[nonprofits$EIN == "721562468", c("EIN", "name", "state", "zip5")]

```
