---
title: "Getting Started"
author: "Olivia Beck"
date: 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(compensator)
```

**compensator** is a package to generate CEO compensation appraisals for nonprofits. 


Add brief paragraph about why we built this package. 

Every appraisal will have three steps. 
1. User provides basic information about the nonprofit they want to obtain an appraisal for. 
2. User provides characteristics of organizations they want to compare the input nonprofit to. 
3. Returns an appraisal for CEO compensation for it input nonprofit. 


## A Few Simple Examples

### Example 1

Say Nonprofit X wants an appraisal for their CEO compensation. Nonprofit X is a shelter for women who have experienced domestic violence and for their children located in Kansas City, MO. Their annual expenses are approximately $5,650,0000. 

**Step 1:** The user defines basic information about their organization using the `get_org_values()` function. See the documentation of `get_org_values()` for detailed information about the arguments. 


```{r}
orgX <- get_org_values(state = "MO",
                      location.type = "metro",
                      total.expense = 5650000,
                      ntee = "P43")
```


**Step 2:** The user defines characteristics about the organizations they want to compare Nonprofit A to in a list with items: 

* `type.org`: vector of the types of organization you want to include. Options are  RG, AA, MT, PA, RP, MS, MM, and/or NS. 
* `broad.category`: vector of broad categories you wish to include in returned data set Options are ART, EDU, ENV, HEL, HMS, IFA, PSB, REL, MMB, UNU, UNI, and/or HOS 
* `major.group vector`: of major groups you wish to include in returned data set. Options are A-Z.
* `division`: vector of divisions you wish to include. Divisions exist entirely inside major groups. We suggest you do not use this parameter if you have more than one item in `major.group`. Options are 0, 2, 3, ..., 9 (1 is not an option. 
* `subdivision`: vector of subdivision you wish to include. Subdivisions exist entirely inside divisions. We suggest you do not use this parameter if you have more than one item in `division`. Options are 0 - 9. 
* `univ`: TRUE of FALSE. Are universities to be included?
* `hosp`: TRUE of FALSE, Are hospitals to be included?
* `location.type`: vector of "metro", "suburban", "town", and/or "rural" for which city types to include
* `state`: vector of 2 letter state abbreviations to be included
* `total.expense`: vector of c(min,max) of range of total expenses to be included

See ___ vignette for detailed information about these options. 

For our example, say the user is only interested in comparing themselves to other nonprofits involve in health care, mental health & crisis intervention, voluntary health associations & medical disciplines, crime & legal-related, and general human services. They also only want to consider other nonprofits with total expenses between \$1,000,000 and \$10,000,000 located in Missouri, Kansas, Oklahoma, Arkansas, Iowa, and Illinois. 


```{r}
search.criteriaX <-
  list(
    type.org = "RG",
    broad.category = c("HEL", "HMS"),
    major.group =  c("E", "F", "G", "I", "P"),
    division = NA,
    subdivision = NA,
    univ = FALSE,
    hosp = FALSE,
    location.type = NA,
    state = c("MO", "KS", "OK", "AR", "IA", "IL"),
    total.expense = c(1000000 , 10000000)
  )
```

**Step 3:** The user inputs their `org` and `search.criteria` lists into the `get_apprasial()` function. The result is a list with three items: 

* `suggested.range` is the final appraisal. This is the range of salaries that we believe are reasonable for the CEO of Nonprofit X accounting for the CEO pay of all other nonprofits that match the `search.criteria`. 
* `suggested.salary` is the weighted average of CEO compensations for all organizations that matched the search criteria, inversely proportional to the similarity between the given organization and Nonprofit X. 
* `reference.set` is the set of all nonprofits that match `search.critera` and their respective attributes. 

See documentation for `get_apprasial()` for more information about these values. 

See ___ Vignette for detailed information about generating `suggested.range` and `suggested.salary`.  

```{r}
appraisalX <- get_appraisal(orgX, search.criteriaX) 
appraisalX$suggested.range
```

We conclude that a reasonable salary range for a CEO of Nonprofit X is `r dollarize(appraisalX$suggested.range[1])` to `r dollarize(appraisalX$suggested.range[2])`.


### Example 2


**Step 1:** Say Nonprofit Y is a wildlife-preservation nonprofit located in Elko County, Nevada with annual expenses of approximately $1,575,000. 

```{r}
orgY <- get_org_values(state = "NV",
                      location.type = "rural",
                      total.expense = 1575000,
                      ntee = "D30")

```

**Step 2:*** They would like to compare themselves to all other animal related nonprofits located not located in metropolitan areas with annual expenses less that $10,000,000. 

```{r}
search.criteriaY <-
  list(
    type.org = "RG",
    broad.category = c("ENV"),
    major.group =  c("D"),
    division = NA,
    subdivision = NA,
    univ = FALSE,
    hosp = FALSE,
    location.type = c("rural", "town", "suburban"),
    state = NA,
    total.expense = c(0 , 10000000)
  )
```

**Step 3:** They generate their appraisal: 


```{r}
appraisalY <- get_appraisal(orgY, search.criteriaY) 
appraisalY$suggested.range
```

We conclude that a reasonable salary range for a CEO of Nonprofit Y is `r dollarize(appraisalY$suggested.range[1])` to `r dollarize(appraisalY$suggested.range[2])`.


### Example 3 

**Step 1:* Say Nonprofit Z is a nonprofit dedicated to fundraising for other nonprofits dedicated to the treatment and prevention of cancer. They are located in Washington D.C. with annual expense of $13,000,000. 

```{r}
orgZ <- get_org_values(state = "DC",
                      location.type = "metro",
                      total.expense = 13000000,
                      ntee = "G1230")

```

**Step 2:* They would like to compare themselves to other fundraising nonprofits regardless of what type organizations they provide support to. 

```{r}
search.criteriaZ <-
  list(
    type.org = c("MS", "MM", "NS"),
    broad.category = NA,
    major.group =  NA,
    division = NA,
    subdivision = NA,
    univ = FALSE,
    hosp = FALSE,
    location.type = NA,
    state = NA,
    total.expense = c(0 , Inf)
  )
```


**Step 3:* They generate their appraisal: 


```{r , warning=F}
appraisalZ <- get_appraisal(orgZ, search.criteriaZ) 
appraisalZ$suggested.range
```

We conclude that a reasonable salary range for a CEO of Nonprofit Z is `r dollarize(appraisalZ$suggested.range[1])` to `r dollarize(appraisalZ$suggested.range[2])`.
